\def\DevnagVersion{2.13}% Assignment 2 for CMSC723
% Finite State Machines

\documentclass[11pt]{article}
\usepackage{latexsym}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[usenames,dvipsnames]{color}
\usepackage[pdftex]{graphicx}


%@modernhindi

\hypersetup{colorlinks=true,linkcolor=blue}

% \date{}
\begin{document}
\begin{center}
{\Large{\textbf{ Natural Language Processing:  }}}\\
\mbox{}\\
{\Large{Assignment 5: Qu' bopbe' paqvam}}\\
% \mbox{}\\
% (\textsc{Not For Credit})\\
\mbox{}\\
{\large{Jordan Boyd-Graber}}\\
\mbox{}\\
{\large{Out: \textbf{16. September 2014}\\Due: \textbf{10. October 2014}}}\\
\end{center}


% \maketitle
\lstset{stringstyle=\ttfamily,language=Python,showstringspaces=False,tabsize=8,frameround=tttt,
		,keywordstyle=\color{Orange}\bfseries, stringstyle=\ttfamily\color{Green}
		,columns=fullflexible,identifierstyle=\ttfamily
		% , commentstyle=\itshape\color{Red}
}

\section*{Introduction} % (fold)
\label{sec:introduction}
As always, check out the Github repository with the course homework templates:

\url{git://github.com/ezubaric/cl1-hw.git}

The code for this homework is in the \texttt{hw5} directory.

\section{Tagging and Tag Sets (10 points)}

\subsection{When taggers go bad (5 points)}

Consider the following sentences:
\begin{enumerate}
\item British Left Waffles on Falkland Islands
\item Teacher Strikes Idle Kids
\item Clinton Wins Budget; More Lies Ahead
\item Juvenile Court to Try Shooting Defendant
\end{enumerate}

Choose one of these sentences and tag it in
two different (but plausible) ways.


\subsection{Exploring the tag set (5 points)}
There are 265 distinct words in the Brown Corpus having exactly four possible tags (assuming nothing is done to normalize the word forms).
\begin{enumerate}
\item Create a table with the integers 1\dots 10 in one column, and the number of distinct
words in the corpus having $\{1, \dots, 10\}$ distinct tags.
\item For the word with the greatest number of distinct tags, print out sentences from
the corpus containing the word, one for each possible tag.

\end{enumerate}

\section{Viterbi Algorithm (30 Points)}

Consider the following sentences written in Klingon.  For each
sentence, the part of speech of each ``word'' has been given (for ease
of translation, some prefixes/suffixes have been treated as words),
along with a translation.  Using these training sentences, we're going
to build a hidden Markov model to predict the part of speech of an
unknown sentence using the Viterbi algorithm.  Do not use log probabilities (you can later, though).

\begin{tabular}{llllll}
  N  & PRO  & V   & N         & PRO  \\
pa'Daq   & ghah & taH & tera'ngan & 'e   \\
room (inside)    & he   & is  & human     & of   \\
\multicolumn{6}{c}{{\em The human is in the room}} \\
\end{tabular}

\begin{tabular}{llll}
V   & N      & V    & N              \\
ja'chuqmeH & rojHom & neH  & tera'ngan        \\
in order to parley  & truce  & want & human \\
\multicolumn{4}{c}{{\em The enemy commander wants a truce in order to parley}}
\end{tabular}

\begin{tabular}{lllllll}
N         & V      & N     & CONJ & N         & V    & N \\
tera'ngan & qIp   & puq   & 'eg  & puq & qIp  & tera'ngan \\
human     & bit  & child & and  & child & bit & child \\
\multicolumn{7}{c}{{\em The child bit the human, and the human bit the child}}
\end{tabular}

\subsection{Emission Probability (10 points)}

Compute the frequencies of each part of speech in the table below for
nouns and verbs.  We'll use a smoothing factor of 0.1 (as discussed in
class) to make sure that no event is impossible; add this number to
all of your observations.  Two parts of speech have already been
done for you.  After you've done this, compute the emission probabilities in a similar table.

\begin{center}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
   & NOUN & VERB & CONJ & PRO \\
\hline
'e &   &   &  0.1    &  1.1   \\
\hline
'eg &     &     &  1.1    &  0.1 \\
\hline
ghaH &     &    &  0.1    &  1.1  \\
\hline
ja'chuqmeH &     &     &  0.1    &  0.1   \\
\hline
legh &     &     &  0.1    &  0.1 \\
\hline
neH &     &     &  0.1    &  0.1   \\
\hline
pa'Daq &     &     &  0.1    &  0.1 \\
\hline
puq &     &     &  0.1    &  0.1 \\
\hline
qIp &     &     &  0.1    &  0.1 \\
\hline
rojHom &     &     &  0.1    &  0.1\\
\hline
taH &     &     &  0.1    &  0.1 \\
\hline
tera'ngan &     &     &  0.1    &  0.1 \\
\hline
yaS &     &     &  0.1    &  0.1   \\
\hline
\end{tabular}
\end{center}

\subsection{Start and Transition Probability (5 points)}

Now, for each part of speech, total the number of times it
transitioned to each other part of speech.  Again, use a smoothing
factor of 0.1.  After you've done this, compute the start and transition probabilities.

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
      & NOUN & VERB & CONJ & PRO \\
\hline
START &  &   &  &   \\
\hline
N     &   &   & 1.1  & 2.1  \\
\hline
V     &   &   & 0.1  & 0.1  \\
\hline
CONJ  &   &   &  0.1 & 0.1 \\
\hline
PRO   &  &   &  0.1 & 0.1 \\
\hline
\end{tabular}
\end{center}

\subsection{Viterbi Decoding (15 points)}

Now consider the following sentence: ``tera'ngan legh yaS''.

\begin{enumerate}

\item Compute the probability of the sequence \textsc{noun}, \textsc{verb}, \textsc{noun}.  

\item Create the decoding matrix of this sentence $\ln \delta_{n}(z)$ (word positions are columns, rows are parts of speech).  Only provide log probabilities, and only use base 2.

\begin{center}
\begin{tabular}{l|ccc}
	POS & $n=1$ & $n=2$ & $n=3$ \\
	\hline
	$z=$\textsc{n} & & & \\
	$z=$\textsc{v} & & & \\
	$z=$\textsc{conj} & & & \\	
	$z=$\textsc{pro} & & & \\	
	\hline
\end{tabular}
\end{center}

\item What is the most likely sequence of parts of speech?

\item Let's compare this to the probability of your previous answer.
	\begin{enumerate}
		\item How does this compare to the sequence \textsc{noun}, \textsc{verb}, \textsc{noun}?
		\item Which is more plausible linguistically?
		\item Does an \textsc{hmm} model encode the intuition that you used to answer the previous question?
	\end{enumerate}
\item (For fun, not for credit) What do you think this sentence means?  What word is the subject of the sentence?

\end{enumerate}



\end{document}
